---
title: "Fuentes de información"
output:
  html_notebook:
    toc: true
    toc_float:
      smooth_scroll: false
      collapsed: false
    df_print: paged
---

# Packages

```{r}
library(tidyverse)
library(lubridate)
library(patchwork)
library(raster)
library(sf)
library(rgdal)
library(gdalUtils)

# Theme for ggplot2
theme_set(theme_minimal())
```

# Join data

```{r, message=FALSE, warning=FALSE}
# Initial data
data_kyoto <- read_csv("../data/raw/kyoto.csv") %>% 
  mutate(new_location = "Kyoto")
data_liestal <- read_csv("../data/raw/liestal.csv")%>% 
  mutate(new_location = "Liestal")
data_wdc <- read_csv("../data/raw/washingtondc.csv")%>% 
  mutate(new_location = "Washington")

# Others
data_japan <- read_csv("../data/raw/japan.csv")%>% 
  mutate(new_location = "Japan")
data_metwiss <- read_csv("../data/raw/meteoswiss.csv")%>% 
  mutate(new_location = "MeteoSwiss")
data_skorea <- read_csv("../data/raw/south_korea.csv")%>% 
  mutate(new_location = "South Korea")

# Join
full_data <-
  bind_rows(data_kyoto,
            data_liestal,
            data_wdc,
            data_japan,
            data_metwiss,
            data_skorea)
```

# Export data

```{r}
write_csv(full_data, file =  "../data/processed/full_data_initial.csv")
```

# Exploratory Data Analysis

```{r}
full_data <- read_csv("../data/processed/full_data_initial.csv")
```

## Distributions

```{r, fig.align="center"}
full_data %>%
  ggplot(aes(
    x = new_location,
    y = bloom_doy,
    fill = new_location,
    color = new_location
  )) +
  geom_boxplot(alpha = 0.5, show.legend = FALSE)
```

## Monthly proportion 

```{r}
full_data %>%
  mutate(bloom_date = as.Date(bloom_date),
         month = month(bloom_date)) %>%
  count(month, new_location) %>%
  group_by(month) %>%
  mutate(total_month = sum(n)) %>%
  ungroup() %>%
  mutate(proportion = n / total_month) %>%
  ggplot(aes(
    x = month,
    y = proportion,
    fill = new_location,
    color = new_location
  )) +
  geom_col(alpha = 0.7) +
  coord_flip() +
  scale_x_continuous(breaks = seq(1, 12, 1))
```

## Monthly frequence 

```{r}
full_data %>%
  mutate(bloom_date = as.Date(bloom_date),
         month = month(bloom_date)) %>%
  count(month, new_location) %>%
  group_by(month) %>%
  mutate(total_month = sum(n)) %>%
  ggplot(aes(
    x = month,
    y = total_month,
    fill = new_location,
    color = new_location
  )) +
  facet_wrap( ~ new_location, scales = "free") +
  geom_col(alpha = 0.7, show.legend = FALSE) +
  scale_x_continuous(breaks = seq(1, 12, 1))
```

## Historical time series 

```{r}
full_data %>%
  mutate(bloom_date = as.Date(bloom_date)) %>%
  group_by(bloom_date, new_location) %>%
  summarise(median_bloom = median(bloom_doy, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(aes(x = bloom_date, y = median_bloom)) +
  facet_wrap(~ new_location, scales = "free") +
  geom_line(alpha = 0.5) +
  geom_smooth()
  
```

## Time series: last century

```{r}
full_data %>%
  mutate(bloom_date = as.Date(bloom_date)) %>%
  filter(bloom_date >= "1900-01-01") %>% 
  group_by(bloom_date, new_location) %>%
  summarise(median_bloom = median(bloom_doy, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(aes(x = bloom_date, y = median_bloom)) +
  facet_wrap(~ new_location, scales = "free") +
  geom_line(alpha = 0.5) +
  geom_smooth()
```

## Scatter plot

```{r}
full_data %>% 
  select(new_location, lat, long, bloom_doy) %>% 
  pivot_longer(cols = c(lat, long)) %>% 
  ggplot(aes(x = value, y = bloom_doy)) +
  facet_wrap(~new_location~name, scales = "free") +
  geom_point(alpha = 0.05) +
  geom_smooth()
```

## Japan

```{r, fig.width=8, fig.height=4}
g1 <- full_data %>% 
  filter(new_location == "Japan") %>% 
  ggplot(aes(x = long, y = lat, z = bloom_doy)) +
  stat_summary_hex(alpha = 0.8, bins = 40, fun = "mean") +
  scale_fill_viridis_c() +
  labs(fill = "mean") +
  theme(legend.position = "top")

g2 <- full_data %>% 
  filter(new_location == "Japan") %>% 
  ggplot(aes(x = long, y = lat, z = bloom_doy)) +
  stat_summary_hex(alpha = 0.8, bins = 40, fun = "median") +
  scale_fill_viridis_c() +
  labs(fill = "median") +
  theme(legend.position = "top")

g3 <- full_data %>% 
  filter(new_location == "Japan") %>% 
  ggplot(aes(x = long, y = lat, z = bloom_doy)) +
  stat_summary_hex(alpha = 0.8, bins = 40, fun = "sd") +
  scale_fill_viridis_c() +
  labs(fill = "sd") +
  theme(legend.position = "top")

g1 + g2 + g3
```

## Mapa Japan

```{r}
map_japan <- getData(name = "GADM", country = "JPN", level = 0,
                     path = "../data/processed/")
map_sf <- map_japan %>% st_as_sf()
map_sf %>% 
  ggplot() + 
  geom_sf() +
  geom_point(aes(x = long, y = lat, color = bloom_doy), alpha = 0.8,
             data = full_data %>% filter(new_location == "Japan"),
             size = 1)  +
  scale_color_viridis_c() 
```

# Data Worldclim

## Data

- [Description Bioclim](https://worldclim.org/data/bioclim.html)
- [Data format](https://worldclim.org/data/v1.4/formats.html)

```{r}
bioclim_data <- getData("worldclim", var = "bio", res = 2.5, 
                        path = "../data/processed/")

cherry_sf <- full_data %>% 
  st_as_sf(coords = c(3, 2),
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

cherry_bioclim <- extract(bioclim_data, cherry_sf) %>% 
  as_tibble() %>% 
  bind_cols(full_data, .)
```

## Exploratory Data Analysis

### Scatter plot 1

```{r, fig.width=8, fig.height=6}
cherry_bioclim %>% 
  dplyr::select(bloom_doy, starts_with("bio")) %>% 
  pivot_longer(cols = -bloom_doy) %>% 
  ggplot(aes(x = value, y = bloom_doy)) +
  facet_wrap(~name, scales = "free") +
  geom_point(alpha = 0.01) +
  geom_smooth()
```

### Scatter plot 2

```{r, fig.width=8, fig.height=6}
cherry_bioclim %>% 
  dplyr::select(bloom_doy, new_location, starts_with("bio")) %>% 
  pivot_longer(cols = -c(bloom_doy, new_location)) %>% 
  ggplot(aes(x = value, y = bloom_doy, color = new_location)) +
  facet_wrap(~name, scales = "free") +
  geom_point(alpha = 0.01) +
  geom_smooth(se = FALSE)
```

# Data SoilGrids

- [Variable description](https://www.isric.org/explore/soilgrids/faq-soilgrids#What_is_SoilGrids)
- [Resources SoilGrids](https://files.isric.org/soilgrids/latest/data/)
- [SoilGrids from R](https://git.wur.nl/isric/soilgrids/soilgrids.notebooks/-/blob/master/markdown/webdav_from_R.md)

## Tiff

```{r}
# Soil components 
bulk_density <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/bdod/bdod_15-30cm_mean.vrt"
caption_exchange <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/cec/cec_15-30cm_mean.vrt"
vol_coarse <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/cfvo/cfvo_15-30cm_mean.vrt"
clay <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/clay/clay_15-30cm_mean.vrt"
nitro <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/nitrogen/nitrogen_15-30cm_mean.vrt"
ph_soil <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/phh2o/phh2o_15-30cm_mean.vrt"
sand <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/sand/sand_15-30cm_mean.vrt"
silt <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/silt/silt_15-30cm_mean.vrt"
soc <-"/vsicurl/https://files.isric.org/soilgrids/latest/data/soc/soc_15-30cm_mean.vrt"
ocd <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/ocd/ocd_15-30cm_mean.vrt"
ocs <- "/vsicurl/https://files.isric.org/soilgrids/latest/data/ocs/ocs_0-30cm_mean.vrt"

# list
list_soil <- list(bulk_density, caption_exchange, vol_coarse, clay, nitro,
                  ph_soil, sand, silt, soc, ocd, ocs)

# Functions
source("../code-R/functions/limits_function.R")
source("../code-R/functions/names_tif.R")
source("../code-R/functions/generator_tif.R")
source("../code-R/functions/all_tif.R")

# Vector of location
list_location <- cherry_sf %>%
  pull(new_location) %>% 
  as_factor() %>% 
  levels()

# Generation tif
all_tif(x = list_location, data_sf = cherry_sf, list_vrt = list_soil)
```

## Data

```{r}

```


```{r}

# Import data
japon_tif <- raster("../data/processed/prueba_japon.tif") 
nuevo_sistema <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"
prueba_nueva <- projectRaster(japon_tif, crs = nuevo_sistema)


cherry_bioclim <- extract(prueba_nueva, prueba) %>% 
  as_tibble() %>% 
  bind_cols(full_data, .)
```


